---
title: "Exam 2"
format: html
---

Load Data
```{r}
library(tidytuesdayR)
library(tidyverse)
library(rnaturalearth)
library(sf)
```
Load Data
```{r}
tuesdata <- tt_load('2020-02-18')
fc <- tuesdata$food_consumption
```
Inspect Data
```{r}
str(fc)
```

```{r}
head(fc,22)
```
```{r}
fc_clean <- fc %>%
  mutate(food_category = case_when(
    food_category == "Lamb & Goat" ~ "Lamb",
    food_category == "Milk - inc. cheese" ~ "Dairy",
    food_category == "Wheat and Wheat Products" ~ "Wheat",
    food_category == "Nuts inc. Peanut Butter" ~ "Nuts",
    TRUE ~ food_category  
  ))

```

```{r}
head(fc_clean,22)
```

Research Question 1

Which five countries consume the most food are?
-show the answer as a 2 column table and as a gragh
-sort the countries based on consumption from largest to smallest
-Use the cleaned datafrme
```{r}
top5_consumers <- fc_clean %>%
  group_by(country) %>%
  summarize(total_consumption = sum(consumption, na.rm = TRUE)) %>%
  arrange(desc(total_consumption)) %>%
  slice_head(n = 5)

top5_consumers

```
```{r}
ggplot(top5_consumers, aes(x = reorder(country, -total_consumption), y = total_consumption, fill = country)) +
  geom_col() +
  labs(
    title = "Top 5 Countries by Total Food Consumption",
    x = "Country",
    y = "Total Consumption (kg/person/year)",
    caption = "Source: TidyTuesday 2020-02-18. 
    Based on total per capita consumption summed across all food categories."
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```

Which top 5 countries consume each food are?
- show the answer as a 3-column table(show the food category as the first column follwed by country and consumption) and as a gragh
-For each food, sort the countries based on consumption from largest to smallest
-Use the cleaned dataframe

```{r}
top5_each_food <- fc_clean %>%
  group_by(food_category) %>%
  slice_max(order_by = consumption, n = 5) %>%
  ungroup()
top5_each_food <- fc_clean %>%
  group_by(food_category, country) %>%
  summarize(consumption = sum(consumption, na.rm = TRUE), .groups = "drop") %>%
  group_by(food_category) %>%
  slice_max(order_by = consumption, n = 5) %>%
  ungroup()
top5_each_food

```

```{r}

library(forcats)

n_pages <- length(unique(top5_each_food$food_category))

for (i in 1:n_pages) {
  p <- ggplot(top5_each_food, aes(x = fct_reorder(country, consumption), y = consumption, fill = country)) +
    geom_col(show.legend = FALSE) +
    coord_flip() +
    ggforce::facet_wrap_paginate(~ food_category, scales = "free_y", ncol = 1, nrow = 1, page = i) +
    labs(
      title = "Top 5 Countries Consuming Each Food Category",
      x = "Country",
      y = "Consumption (kg/person/year)",
      caption = paste("Source: TidyTuesday 2020-02-18. Page", i, "of", n_pages)
    ) +
    theme_minimal(base_size = 11) +
    theme(
      strip.text = element_text(size = 11, face = "bold"),
      plot.caption = element_text(hjust = 0, size = 8, color = "gray40")
    )
  
  print(p)  
}



```
Research Question

What does the consumption of each food look like?
-show a chropleth map for each food.
-use the cleaned dataframe
```{r}
ne_countries(returnclass = "sf") |>
  select(name, geometry) |>
  mutate(name = ifelse(name == "United States of America", "USA", name)) |>
  mutate(name = ifelse(name == "Bosnia and Herz.", "Bosnia and Herzegovina", name)) |>
  mutate(name = ifelse(name == "Czechia", "Czech Republic", name)) |>
  mutate(name = ifelse(name == "Taiwan", "Taiwan. ROC", name)) |>
  left_join(
    fc_clean |>
      select(-co2_emmission) |>  
      group_by(food_category) |>
      mutate(consumption = (consumption - mean(consumption)) / sd(consumption)) |> 
      ungroup() |>
      pivot_wider(names_from = food_category, values_from = consumption), 
    by = join_by(name == country)  
  ) |>
  pivot_longer(cols = -c(name, geometry), names_to = "food_category", values_to = "consumption") |>  
  ggplot() +
  labs(
    title = "Standardized Food Consumption by Country and Category",
    fill = "Consumption (z-score)",
    caption = "Source: TidyTuesday 2020-02-18. Consumption values standardized (z-score) by food category.\nMap shows relative per capita food consumption across countries."
  )+
  geom_sf(aes(fill = consumption)) +
  facet_wrap(~ food_category) +
  theme(legend.position = "bottom")

```

